### 2015-01-06 postscript.tex

\section{Postscript}

We began this work in late 2013, generating a reduced package for \emph{TauRoast} based
on the configuration of a standard machine at Notre Dame at the time.  In the course
of writing this paper in spring 2014, almost everything about the computing environment
changed: the operating system was upgraded, a new version of CMSSW was released, and
our local HEP users switched from using CVS to CVMFS for accessing CMSSW.
The original script provided by the author failed to run on the same machine.
But, the reduced package we created three months earlier in the old environment
worked just fine on the new machine.